---
title: "to_go_alcohol_policy_analyses"
author: "Drew Walker"
date: '2022-07-12'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(measurements)
library(units)
library(table1)
library(nlme)
library(here)
library(psych)
library(misty)
here()
library(table1)
library(ruca)
library(lme4)
library(sjPlot)
library(sjmisc)
library(patchwork)
library(sjlabelled)
#install.packages('TMB', type = 'source')
library(TMB)
```

For this analysis, we will be utilizing data on alcohol to-go policies for bar and restaurant outlets located within 5 miles of college campuses in the US during the COVID-19 pandemic.

We will utilize multilevel models to account for census tract clustering of outcomes, while assessing the associations between:

1.  State to-go policy (predictor) and bar/restaurant beer and vodka prices (outcomes)

2.   The state policy on whether bar/restaurants could sell to-go beer/wine and mixed drinks (predictor) with whether bars and restaurants around college campuses actually sold to-go beer, wine, and mixed drinks (outcomes).

We present first a set of unadjusted models between each policy and bar/restaurant outcome clustered at the census tract level, followed by adjusted models with the inclusion of the following set of covariates:

-   Distance from school (natural log)

-   Diversity score

-   Disadvantage score

```{r load-data, include=FALSE}
combined_covariates <- readRDS("census_tracts_bar_policy_togo_covariates.rds")
combined_covariates2 <- combined_covariates %>% 
  unnest(census_data) %>% 
  select(UID,geographies.Census.Tracts.GEOID,geographies.Census.Tracts.OID,geographies.Census.Tracts.TRACT,addressComponents.zip)

clean_combined_covariates <- left_join(combined_covariates,combined_covariates2,by="UID")

# Getting categorization of restaurant or bar
# Rules from team: 
# Restaurant: open midnights <= 3 nights a week
# Bar = open midnights >3 times a week 
# bar_vs_rest : 1 is bar, 0 is rest
clean_combined_coavriates_rest_bar <- clean_combined_covariates %>% 
  mutate(bar_vs_rest = case_when(X.NightsOpenPast12amEachWeek >= 3 ~ 1,
                                 X.NightsOpenPast12amEachWeek < 3 ~ 0))



  
clean_combined_covariates_trim <- clean_combined_coavriates_rest_bar %>% 
  select(UID,SID=SID.x, SchoolName = SchoolName.x,
         BarName = BarName.x,
         Suggested_code_bar,Suggested_code_rest,NABCA_spirits_bars,NABCA_spirits_rest,geographies.Census.Tracts.GEOID,geographies.Census.Tracts.OID,geographies.Census.Tracts.TRACT,addressComponents.zip,alcohol_togo,FIPS,bar_vs_rest)

wave_2_data <- read_csv("Results for price stuff/Wave2_cleaned_dw.csv")
wave_2_data_merge <- wave_2_data %>% 
  select(UID,SID,RC_Masks_customer,RC_Masks_staff,RC_distance,alcohol_togo, Beer_Price2, Vodka_price2,Cigarettes_inside,Ecig_inside)

wave1_raw <- read_csv("Results for price stuff/Wave1_cleaned.csv")

wave1_clean <- wave1_raw %>% 
  select(UID,SID,Vodka_price,Beer_Price)


both_wave_data_covariates <- left_join(clean_combined_covariates_trim, wave1_clean, by = "UID")
both_wave_data <- left_join(both_wave_data_covariates,wave_2_data_merge, by = "UID")

both_wave_data <- both_wave_data %>% 
  mutate(beer_price_change = Beer_Price2-Beer_Price,
         vodka_price_change = Vodka_price2-Vodka_price) %>% 
  select(-SID.x) %>% 
  rename(alcohol_togo = alcohol_togo.y,
         zip = addressComponents.zip,
         tract_code = geographies.Census.Tracts.GEOID)
## Connect with RUCA codes at zip level
ruca <- read_csv("ruca2010revised.csv",skip = 1)
ruca_clean <- ruca %>% 
  select(tract_code =`State-County-Tract FIPS Code (lookup by address at http://www.ffiec.gov/Geocode/)`,
         ruca = `Primary RUCA Code 2010`,
         secondary_ruca =`Secondary RUCA Code, 2010 (see errata)`
         )
data("classifications")
ruca_clean <- left_join(ruca_clean,classifications, by = "ruca")
#Dataframe with addresses, rucacodes, and record_ids 
addresses_with_ruca_codes <- left_join(both_wave_data, ruca_clean, by = "tract_code")

addresses_with_ruca_codes$ruca <- as.factor(addresses_with_ruca_codes$ruca)
table1(~ruca + description, data = addresses_with_ruca_codes)


# Add Census tract-level covariates from Pamela 
ct_covariates <- read_csv("ct-level covar.csv")
ct_covariates_clean <- ct_covariates %>% 
  rename(tract_code=fips_ct)
ct_covariates_clean$tract_code <- as.character(ct_covariates_clean$tract_code)

addresses_with_ruca_codes1 <- left_join(addresses_with_ruca_codes,ct_covariates_clean, by = "tract_code")


# Adding new policy variable from E-mail 

new_policy_vars <- read_csv("Comparing state COVID Alc Policy Sources PT 2.csv")
new_policy_vars_clean <- new_policy_vars %>%  
  select(FIPS,Suggested_code_bar3, Suggested_code_rest4, Suggested_code_bar_spirits, Suggested_code_rest_spirits)  


addresses_with_ruca_codes1 <- left_join(addresses_with_ruca_codes1,new_policy_vars_clean, by = "FIPS")

addresses_with_ruca_codes2 <- addresses_with_ruca_codes1 %>% 
    mutate(bw_togo_legal = case_when(bar_vs_rest == "1" & Suggested_code_bar3 == "1"|bar_vs_rest == "0" & Suggested_code_rest4 ~ "1",
                                     TRUE ~ "0"),
         spirits_togo_legal = case_when(bar_vs_rest == "1" & Suggested_code_bar_spirits == "1"| bar_vs_rest =="0" & Suggested_code_rest_spirits == "0" ~ "1",
                                    TRUE ~ "0"))

# Merging Beer/Wine/Mixed drink togo sales outcome
beer_mixed_wine_sales_outcomes <- read_csv("merged_clean_bwm_wave2_edited2.csv")

beer_mixed_wine_sales_outcomes_clean <- beer_mixed_wine_sales_outcomes %>% 
  select(UID,beer,wine,mixed_drinks)

addresses_with_ruca_codes3 <- left_join(addresses_with_ruca_codes2,beer_mixed_wine_sales_outcomes_clean, by = "UID")  

distance_data <- read_csv("Wave 1 Data and Sampling Frame 10 26 20.csv")


distance_data_clean <- distance_data %>% 
  mutate(distance_clean = case_when(str_detect(Distance, "ft| ft| feet|feet")~ conv_unit(as.numeric(str_remove(distance_data$Distance, "ft| ft| feet|feet")),"ft","mi"), TRUE ~ as.numeric(Distance))) %>% 
  select(UID, Distance, distance_clean) %>% 
  mutate(std_distance = std(distance_clean),
         log_distance = log(distance_clean))

# Now create or doublecheck whether or not place sold togo Beer Wine / Spirit as an outcome 

addresses_with_ruca_codes <- left_join(addresses_with_ruca_codes3, distance_data_clean, by = "UID")
```

## Exploratory Data Analysis

-   univariates, bivariates of outcomes/predictors and covariates.
-   Reorg this section

```{r eda}
# Outcome descriptive statistics
describe(addresses_with_ruca_codes$distance_clean)
describe(both_wave_data$vodka_price_change)
describe(both_wave_data$beer_price_change)
hist(both_wave_data$vodka_price_change)
hist(both_wave_data$beer_price_change)
# Should we remove $17 beer data point?

## PREDICTORS
both_wave_data$Suggested_code_bar <- as.factor(both_wave_data$Suggested_code_bar)
both_wave_data$Suggested_code_rest <- as.factor(both_wave_data$Suggested_code_rest)
both_wave_data$NABCA_spirits_bars <- as.factor(both_wave_data$NABCA_spirits_bars)
both_wave_data$NABCA_spirits_rest <- as.factor(both_wave_data$NABCA_spirits_rest)
both_wave_data$alcohol_togo <- as.factor(both_wave_data$alcohol_togo)

#Covariates 
table1(~Suggested_code_bar + Suggested_code_rest + NABCA_spirits_bars + NABCA_spirits_rest + alcohol_togo + beer + mixed_drinks + wine, data=addresses_with_ruca_codes)
# Disadvantage
# Distance
# Diversity 
```

-   224 sites with complete alcohol to-go variable (179 missing, 44.4% of 403)

-   Bars/restaurants nested within census tracts and states

-   Policies of states and practices of local areas around alcohol sales are theoretically thought to influence prices

# 3. Model Level Structure

-   **Outcome**s: Price of vodka and price of beer

-   **Level 1 Predictors**:

    -   Bar or restaurant to-go alcohol practice

-   **Level 2 Predictors**:

    -   Census tract (disadvantage

-   **Level 3 Predictor**:

    -   State alcohol to-go APIS/NABCA policies (allowing or not allowing sales of alcohol)

# Build Null Model (Outcomes only)

-   In this model, we'll include the outcome (beer and vodka price changes) and test the clustering effects at tract code with clustering effects at zip code level.

```{r build-null}
# lmer version
beer_null = lmer(beer_price_change ∼ (1|tract_code), both_wave_data, REML=FALSE)
summary(beer_null)

vodka_null = lmer(vodka_price_change ∼ (1|tract_code), both_wave_data, REML=FALSE)


summary(vodka_null)

zip_beer_null = lmer(beer_price_change ∼ (1|zip), both_wave_data, REML=FALSE)
summary(zip_beer_null)

zip_vodka_null = lmer(vodka_price_change ∼ (1|zip), both_wave_data, REML=FALSE)


summary(zip_vodka_null)


tab_model(beer_null,vodka_null, zip_beer_null, zip_vodka_null, dv.labels = c("Null Beer Model", "Null Vodka Model", "Null Beer Zip Model", "Null Vodka Zip Model"), show.aic = TRUE, show.dev = TRUE)
```

# ICC calculations

<https://cran.r-project.org/web/packages/misty/misty.pdf>

## Census tract level for beer/vodka price change

```{r, census-ICCs}
# Beer price change outcome, census tract group
multilevel.icc(both_wave_data$beer_price_change, cluster = both_wave_data$tract_code)

# Vodka price change outcome, census tract group

multilevel.icc(both_wave_data$vodka_price_change, cluster = both_wave_data$tract_code)
# Actually finding very high clustering for vodka price change 

```

## State level for beer/vodka price change

```{r, state-ICCs}
# Beer price change outcome, by state
multilevel.icc(both_wave_data$beer_price_change, cluster = both_wave_data$FIPS)

# Vodka price change outcome, by state 

multilevel.icc(both_wave_data$vodka_price_change, cluster = both_wave_data$FIPS)
# 5% ICC

```
# Beer and Vodka Price Change Models

## Unadjusted beer/togo sale state legality on price 
```{r unadjusted-policy-price-changes}
#Beer and alcohol togo
addresses_with_ruca_codes$bw_togo_legal
unadjusted_beer_price = lmer(beer_price_change ∼ bw_togo_legal + (1|tract_code), addresses_with_ruca_codes, REML=FALSE)

#Vodka price and alcohol togo
unadjusted_vodka_price = lmer(vodka_price_change ∼ spirits_togo_legal + (1|tract_code), addresses_with_ruca_codes, REML=FALSE)

tab_model(unadjusted_beer_price,unadjusted_vodka_price, dv.labels = c("Unadjusted Beer/Wine Togo State Legality on Beer Price ", "Unadjusted Spirits Togo State Legality on Vodka Price"), show.aic = TRUE, show.dev = TRUE)

```

## Adjusted beer/togo sale state legality on price
### Covariates: 

-   Distance from school (natural log)

-   Diversity score

-   Disadvantage score

```{r adjusted-policy-price-changes}
#Beer and alcohol togo

unadjusted_beer_price = lmer(beer_price_change ∼ bw_togo_legal + log_distance + diversity + disadv + (1|tract_code), addresses_with_ruca_codes, REML=FALSE)

#Vodka price and alcohol togo
unadjusted_vodka_price = lmer(vodka_price_change ∼ spirits_togo_legal + log_distance + diversity + disadv + (1|tract_code), addresses_with_ruca_codes, REML=FALSE)

tab_model(unadjusted_beer_price,unadjusted_vodka_price, dv.labels = c("Adjusted Beer/Wine Togo State Legality on Beer Price ", "Adjusted Spirits Togo State Legality on Vodka Price"), show.aic = TRUE, show.dev = TRUE)

```

## Unadjusted togo site practice on price 

```{r unadjusted-price-changes}
#Beer and alcohol togo
addresses_with_ruca_codes$mixed_drinks
unadjusted_beer_price = lmer(beer_price_change ∼ beer + (1|tract_code), addresses_with_ruca_codes, REML=FALSE)

#Vodka price and alcohol togo
unadjusted_vodka_price = lmer(vodka_price_change ∼ mixed_drinks + (1|tract_code), addresses_with_ruca_codes, REML=FALSE)

tab_model(unadjusted_beer_price,unadjusted_vodka_price, dv.labels = c("Unadjusted Bar/Rest Practice of Selling Beer Togo on Beer Price ", "Unadjusted Bar/Rest Practice of Selling Mixed Drinks Togo on Vodka Price"), show.aic = TRUE, show.dev = TRUE)

```

## Adjusted togo site practice on price
### Covariates: 

-   Distance from school (natural log)

-   Diversity score

-   Disadvantage score


```{r adjusted-price-changes}
#Beer and alcohol togo

adjusted_beer_price_practice = lmer(beer_price_change ∼ beer + log_distance + diversity + disadv + (1|tract_code), addresses_with_ruca_codes, REML=FALSE)

#Vodka price and alcohol togo
adjusted_vodka_price_practice = lmer(vodka_price_change ∼ mixed_drinks + log_distance + diversity + disadv + (1|tract_code), addresses_with_ruca_codes, REML=FALSE)

tab_model(adjusted_beer_price_practice,adjusted_vodka_price_practice, dv.labels = c("Adjusted Beer/Wine Togo Site Sales Practice on Beer Price ", "Adjusted Mixed Drinks Togo Site Sales Practice on Vodka Price"), show.aic = TRUE, show.dev = TRUE)

```
# State Policy on Bar/Restaurant Actual Togo Practice Models

## Unadjusted state policy on actual practice sales of togo Beer, Wine and Mixed Drinks

```{r unadjusted-actual-practice}
#Beer/Wine togo legal policy on actual practice of selling beer togo
unadjusted_beer_togo = lmer(beer ∼ bw_togo_legal + (1|tract_code), addresses_with_ruca_codes, REML=FALSE)

#Beer/Wine togo legal policy on actual practice of selling Wine togo

unadjusted_wine_togo = lmer(wine ∼ bw_togo_legal + (1|tract_code), addresses_with_ruca_codes, REML=FALSE)

#Liquor togo legal policy on actual practice of selling mixed drinks togo
unadjusted_mixed_drinks_togo = lmer(mixed_drinks ∼ spirits_togo_legal + (1|tract_code), addresses_with_ruca_codes, REML=FALSE)



tab_model(unadjusted_beer_togo,unadjusted_wine_togo,unadjusted_mixed_drinks_togo, dv.labels = c("Unadjusted Beer/Wine Togo State Policy on Bar/Rest Practice of Selling Beer Togo", "Unadjusted Beer/Wine Togo State Policy on Bar/Rest Practice of Selling Wine Togo", "Unadjusted Spirits Togo State Policy on Bar/Rest Practice of Selling Mixed Drinks Togo"), show.aic = TRUE, show.dev = TRUE)

```

## Adjusted togo site practice on price
### Covariates: 

-   Distance from school (natural log)

-   Diversity score

-   Disadvantage score


```{r adjusted-actual-practice}
#Beer and alcohol togo

adjusted_beer_price_practice = lmer(beer_price_change ∼ beer + log_distance + diversity + disadv + (1|tract_code), addresses_with_ruca_codes, REML=FALSE)



#Beer/Wine togo legal policy on actual practice of selling beer togo
adjusted_beer_togo = lmer(beer ∼ bw_togo_legal + log_distance + diversity + disadv + (1|tract_code), addresses_with_ruca_codes, REML=FALSE)

#Beer/Wine togo legal policy on actual practice of selling Wine togo

adjusted_wine_togo = lmer(wine ∼ bw_togo_legal + log_distance + diversity + disadv + (1|tract_code), addresses_with_ruca_codes, REML=FALSE)

#Liquor togo legal policy on actual practice of selling mixed drinks togo
adjusted_mixed_drinks_togo = lmer(mixed_drinks ∼ spirits_togo_legal + log_distance + diversity + disadv +(1|tract_code), addresses_with_ruca_codes, REML=FALSE)

tab_model(adjusted_beer_togo,adjusted_wine_togo,adjusted_mixed_drinks_togo, dv.labels = c("Adjusted Beer/Wine Togo State Policy on Bar/Rest Practice of Selling Beer Togo", "Adjusted Beer/Wine Togo State Policy on Bar/Rest Practice of Selling Wine Togo", "Adjusted Spirits Togo State Policy on Bar/Rest Practice of Selling Mixed Drinks Togo"), show.aic = TRUE, show.dev = TRUE)
```