---
title: "Job Aid: Multilevel modeling cross-sectional continuous outcomes"
author: "Drew Walker"
date: "3/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(table1)
library(nlme)
library(here)
library(psych)
library(misty)
here()
library(table1)
library(ruca)
library(lme4)
library(sjPlot)
library(sjmisc)
library(patchwork)
library(sjlabelled)
#install.packages('TMB', type = 'source')
library(TMB)
```

# 1. Purpose

-   This job guide is intended to provide a checklist for future reference in building multilevel cross-sectional regression models.

-   For this example, we will be utilizing data on alcohol to-go policies for bar and restaurant outlets located within 5 miles of college campuses in the US during the COVID-19 pandemic.

-   We'll be primarily using R's nlme package

-   <https://cran.r-project.org/web/packages/nlme/nlme.pdf>

# 2. Assumptions

-   Continuous outcomes (though these allow for logistic regression also, incorporation of different logit functions) :

          - price of beer and price of vodka

-   Predictors (binary):

-   Alcohol to-go policy (1st level) (0= did not allow alcohol to go during COVID-19 data collection, 1 = sold alcohol to-go during COVID-19 data collection)

-   Data contained in nested structure of Restaurant or Bars (Lvl 1) nested in Census Tracts (Lvl 2), nested in States (Lvl 3)

-   EDA: see below to assess distributions of outcomes and predictors-- likely will have issues interpreting differences in RUCA codes.

-   Theoretical rationale: For this project, I was asked to lead a regression analysis on bar and restaurant price changes in alcohol around college campuses, in order to determine potential changes in price point around the COVID-19 pandemic as it is associated with to-go alcohol sales. I immediately wondered if this regression should account for nesting within geographic locations, which led to this exploratory analysis used as an example for this guide.

```{r load-data}
combined_covariates <- readRDS("census_tracts_bar_policy_togo_covariates.rds")
combined_covariates2 <- combined_covariates %>% 
  unnest(census_data) %>% 
  select(UID,geographies.Census.Tracts.GEOID,geographies.Census.Tracts.OID,geographies.Census.Tracts.TRACT,addressComponents.zip)

clean_combined_covariates <- left_join(combined_covariates,combined_covariates2,by="UID")
  
clean_combined_covariates_trim <- clean_combined_covariates %>% 
  select(UID,SID=SID.x, SchoolName = SchoolName.x,
         BarName = BarName.x,
         Suggested_code_bar,Suggested_code_rest,NABCA_spirits_bars,NABCA_spirits_rest,geographies.Census.Tracts.GEOID,geographies.Census.Tracts.OID,geographies.Census.Tracts.TRACT,addressComponents.zip,alcohol_togo,FIPS)

wave_2_data <- read_csv("Results for price stuff/Wave2_cleaned_dw.csv")
wave_2_data_merge <- wave_2_data %>% 
  select(UID,SID,RC_Masks_customer,RC_Masks_staff,RC_distance,alcohol_togo, Beer_Price2, Vodka_price2,Cigarettes_inside,Ecig_inside)

wave1_clean <- read_csv("Results for price stuff/Wave1_cleaned.csv")

wave1_clean <- wave1_clean %>% 
  select(UID,SID,Vodka_price,Beer_Price)


both_wave_data_covariates <- left_join(clean_combined_covariates_trim, wave1_clean, by = "UID")
both_wave_data <- left_join(both_wave_data_covariates,wave_2_data_merge, by = "UID")

both_wave_data <- both_wave_data %>% 
  mutate(beer_price_change = Beer_Price2-Beer_Price,
         vodka_price_change = Vodka_price2-Vodka_price) %>% 
  select(-SID.x) %>% 
  rename(alcohol_togo = alcohol_togo.y,
         zip = addressComponents.zip,
         tract_code = geographies.Census.Tracts.GEOID)
## Connect with RUCA codes at zip level
ruca <- read_csv("ruca2010revised.csv",skip = 1)
ruca_clean <- ruca %>% 
  select(tract_code =`State-County-Tract FIPS Code (lookup by address at http://www.ffiec.gov/Geocode/)`,
         ruca = `Primary RUCA Code 2010`,
         secondary_ruca =`Secondary RUCA Code, 2010 (see errata)`
         )
data("classifications")
ruca_clean <- left_join(ruca_clean,classifications, by = "ruca")
#Dataframe with addresses, rucacodes, and record_ids 
addresses_with_ruca_codes <- left_join(both_wave_data, ruca_clean, by = "tract_code")

addresses_with_ruca_codes$ruca <- as.factor(addresses_with_ruca_codes$ruca)
table1(~ruca + description, data = addresses_with_ruca_codes)


# Add Census tract-level covariates from Pamela 
ct_covariates <- read_csv("ct-level covar.csv")
ct_covariates_clean <- ct_covariates %>% 
  rename(tract_code=fips_ct)
ct_covariates_clean$tract_code <- as.character(ct_covariates_clean$tract_code)

addresses_with_ruca_codes <- left_join(addresses_with_ruca_codes,ct_covariates_clean, by = "tract_code")


```

## Exploratory Data Analysis

-   univariates, bivariates of outcomes/predictors and covariates.

```{r eda}
# Outcome descriptive statistics
describe(both_wave_data$vodka_price_change)
describe(both_wave_data$beer_price_change)
hist(both_wave_data$vodka_price_change)
hist(both_wave_data$beer_price_change)
# Should we remove $17 beer data point?

## PREDICTORS
both_wave_data$Suggested_code_bar <- as.factor(both_wave_data$Suggested_code_bar)
both_wave_data$Suggested_code_rest <- as.factor(both_wave_data$Suggested_code_rest)
both_wave_data$NABCA_spirits_bars <- as.factor(both_wave_data$NABCA_spirits_bars)
both_wave_data$NABCA_spirits_rest <- as.factor(both_wave_data$NABCA_spirits_rest)
both_wave_data$alcohol_togo <- as.factor(both_wave_data$alcohol_togo)

#Covariates 
table1(~Suggested_code_bar + Suggested_code_rest + NABCA_spirits_bars + NABCA_spirits_rest + alcohol_togo, data=both_wave_data)

```

-   224 sites with complete alcohol to-go variable (179 missing, 44.4% of 403)

-   Bars/restaurants nested within census tracts and states

-   Policies of states and practices of local areas around alcohol sales are theoretically thought to influence prices

# 3. Model Building

-   **Outcome**s: Price of vodka and price of beer

-   **Level 1 Predictors**:

    -   Bar or restaurant to-go alcohol practice

-   **Level 2 Predictors**:

    -   Census tract

-   **Level 3 Predictor**:

    -   State alcohol to-go APIS/NABCA policies (allowing or not allowing sales of alcohol)

# Build Null Model (Outcomes only)

-   In this model, we'll include the outcome (beer and vodka price changes), as well as the levels, being census tract and state

```{r build-null}
# lmer version
beer_null = lmer(beer_price_change ∼ (1|tract_code), both_wave_data, REML=FALSE)
summary(beer_null)

vodka_null = lmer(vodka_price_change ∼ (1|tract_code), both_wave_data, REML=FALSE)


summary(vodka_null)

zip_beer_null = lmer(beer_price_change ∼ (1|zip), both_wave_data, REML=FALSE)
summary(zip_beer_null)

zip_vodka_null = lmer(vodka_price_change ∼ (1|zip), both_wave_data, REML=FALSE)


summary(zip_vodka_null)


tab_model(beer_null,vodka_null, zip_beer_null, zip_vodka_null, dv.labels = c("Null Beer Model", "Null Vodka Model", "Null Beer Zip Model", "Null Vodka Zip Model"), show.aic = TRUE, show.dev = TRUE)
```

# ICC calculations

<https://cran.r-project.org/web/packages/misty/misty.pdf>

## Census tract level for beer/vodka price change

```{r, census-ICCs}
# Beer price change outcome, census tract group
multilevel.icc(both_wave_data$beer_price_change, cluster = both_wave_data$tract_code)

# Vodka price change outcome, census tract group

multilevel.icc(both_wave_data$vodka_price_change, cluster = both_wave_data$tract_code)
# Actually finding very high clustering for vodka price change 

```

## State level for beer/vodka price change

```{r, state-ICCs}
# Beer price change outcome, by state
multilevel.icc(both_wave_data$beer_price_change, cluster = both_wave_data$FIPS)

# Vodka price change outcome, by state 

multilevel.icc(both_wave_data$vodka_price_change, cluster = both_wave_data$FIPS)
# 5% ICC

```

# Adding Level-1 predictors and level 2 grouping

-   alcohol_togo is a predictor at the site level (level 1) -- whether bar or restaurant allowed sale of to-go alcohol during pandemic , on the outcome of change in beer and vodka price during pandemic

-   tract_code represents the census tract, which is the 2nd level of nesting.

-   For random intercepts (last two mixed models), replace the 1 in the random statement with the variable which you want to allow slopes to vary between groups. See commented code for example (though this example does not converge)

```{r model2-predictors}
#Beer and alcohol togo

lvl1_beer = lmer(beer_price_change ∼ alcohol_togo + (1|tract_code), both_wave_data, REML=FALSE)
summary(lvl1_beer)

#Vodka price and alcohol togo
lvl1_vodka = lmer(vodka_price_change ∼ alcohol_togo + (1|tract_code), both_wave_data, REML=FALSE)
summary(lvl1_vodka)


tab_model(lvl1_beer,lvl1_vodka, dv.labels = c("Beer Price Change Alcohol To-go Predictor Model", "Vodka Price Change Alcohol To-go Predictor Model"), show.aic = TRUE, show.dev = TRUE)


```

# Adding Level-2 predictors

-   Categorical predictor of RUCA code

-   This is just an example, but in fact this is actually not great practice since it's a complicated categorcical variable. Simply add other level variables to model statement, and after \| if desired

-   Adding census-level covariates (from Pamela) of:

    -   **Diversity:** This is Simpson's Diversity Index. It measures the
        probability that two people chosen at random will be from different racial
        and ethnic groups. It is bounded between 0 and 1.

    -   **Disadv:** This is the "disadvantage" index that Deb Furr-Holden
        always uses. It's calculated as ((%college-degree-or-more +
        %owner-occupied) -- (%poverty + %female-headed-HHs))/4. It ranges from -1
        to 1; the negative values are "disadvantaged" and the positive values are
        advantaged. I hate the term disadvantage when ascribed to groups of people
        or places, which is why I keep putting it in quotes BTW.

```{r lvl2-predictors}


#Beer and RUCA code (Census tract lvl)

lvl2_beer = lmer(beer_price_change ∼ diversity + disadv + (1|tract_code), addresses_with_ruca_codes, REML=FALSE)
summary(lvl2_beer)

#Vodka price and RUCA code (census tract lvl)
lvl2_vodka = lmer(vodka_price_change ∼ diversity + disadv + (1|tract_code), addresses_with_ruca_codes, REML=FALSE)
summary(lvl2_vodka)


tab_model(lvl2_beer,lvl2_vodka, dv.labels = c("Beer Ruca Model", "Vodka Ruca Model"), show.aic = TRUE, show.dev = TRUE)





```

No significant effect from RUCA codes or covariates, likely variance is already accounted for by adding random intercept for census-tract level.

# Adding Level-3 predictors and nesting

-   Adding more than 2 levels requires doing random = \~ 1 \| group1/group 2 \*\* This indicates nested structure \*\* <https://stats.stackexchange.com/questions/12768/three-level-hierarchical-regression-using-lmer> \*\* Alternatively, random = \~ 1 \| group1 + random = \~ 1\| group 2 indicates cross-nested structure

```{r state-level-predictors}

lvl3_vodka = lmer(vodka_price_change ∼ alcohol_togo + Suggested_code_bar+(1|tract_code), addresses_with_ruca_codes, REML=FALSE)
summary(lvl3_vodka)

lvl3_beer = lmer(beer_price_change ∼ alcohol_togo + Suggested_code_bar + (1|tract_code), addresses_with_ruca_codes, REML=FALSE)
summary(lvl3_beer)


tab_model(lvl3_beer, lvl3_vodka, dv.labels = c("Beer Price Change State-level Policy Predictors", "Vodka Price Change State-level Policy Predictors"))


togo_answered <- both_wave_data %>% 
  filter(alcohol_togo != "NA") 
sum(is.na(togo_answered$tract_code))


```

# Do state-level policies predict if a bar/restaurant was serving to-go alcohol?

```{r predicting-to-go-policy}

both_wave_data$Suggested_code_bar <- as.factor(both_wave_data$Suggested_code_bar)
both_wave_data$Suggested_code_rest <- as.factor(both_wave_data$Suggested_code_rest)
str(both_wave_data$Suggested_code_bar)
# Does the suggested bar code predict to-go policy at location?
togo_policy_suggested_bar <- glm(formula=alcohol_togo ~ Suggested_code_bar + diversity + disadv, data = addresses_with_ruca_codes, family = "binomial")
summary(togo_policy_suggested_bar)
# Does the suggested restaurant code predict to-go policy at location?

togo_policy_suggested_rest <- glm(formula=alcohol_togo ~ Suggested_code_rest + diversity + disadv, data = addresses_with_ruca_codes, family = "binomial")
summary(togo_policy_suggested_rest)
# Other policy variables?
togo_policy_nabca <- glm(formula=alcohol_togo ~ NABCA_spirits_rest + NABCA_spirits_bars + diversity + disadv, data = addresses_with_ruca_codes, family = "binomial")
summary(togo_policy_nabca)

tab_model(togo_policy_suggested_bar, togo_policy_suggested_rest,togo_policy_nabca, dv.labels = c("Likelihood of Site To-go Practice predicted by State-level Policy Suggested Restaurant Predictor", "Likelihood of Site To-go Practice predicted by State-level Policy Suggested Bar Predictor", "Likelihood of Site To-go practice predicted by NABCA spirits policies"))


```

After running logistic regressions using state-level policy variables (2 suggested codes which researchers ML and PT developed from multiple policy data sources for restaurant and bar policies), and one model using NABCA policy state-level indicators for bars and restaurants. We found no significant association between state-level policies and whether or not a site was serving alcohol to go.

# Display/Interpretation of Results

-   Lvl 1 Variance is displayed via σ2, while 2nd level variance is displayed τ00

```{r tabling}
library(sjPlot)
library(sjmisc)
library(sjlabelled)
tab_model(lvl1_beer, lvl1_vodka, lvl2_beer, lvl2_vodka, lvl3_beer,lvl3_vodka, dv.labels = c("Lvl1 to-go policy on Price of Beer Change", "Lvl1 to-go policy on Price of Vodka Change", "Lvl2 Beer", "Lvl2 Vodka", "Lvl3 Beer with Policy Predictor","Lvl3 Vodka with Policy Predictor"))


```

# Reporting results

-   Interpretation of betas

    Fixed effects betas are discussed similar to standard regression.

    -   ICCs:

        -   We found moderately high ICCs for the clustering of the price changes of vodka at the census tract level (28%-32%), but all other ICCs calculated for State and Census-tract clustering was below 5%.

-   Model fit

    -   To compare model results, we need to look at summary(model) statements individually. In this case, Our 2nd level model, including clustering at the census-tract data level and RUCA codes indicated best fit (AIC= 581). However, as stated previously, this model likely is not meaningful in the interpretation of RUCA code answers due to the limited range of codes in this dataset, likely due to clustering around college campuses.

```{r model-fit}
summary(lvl1_beer) #716.0994	AIC
summary(lvl1_vodka) # 645.6705 AIC
summary(lvl2_beer)  # 581.2696 AIC
summary(lvl3_beer) # 719.3356 AIC
```

## Follow-up analyses

After sending initial results to the team, was asked to perform some additional analyses:

-   Can you run a simple percentage of bars and restaurants (separately) that sell to-go by state policy? I want to understand the last result a little more.

```{r cross-tabs}
# Suggested code policies with site-specific to-go practice 
sjPlot::tab_xtab(var.row = addresses_with_ruca_codes$alcohol_togo, var.col = addresses_with_ruca_codes$Suggested_code_bar, title = "Site practice of Togo sales by Suggested Code Bar", show.row.prc = TRUE)

sjPlot::tab_xtab(var.row = addresses_with_ruca_codes$alcohol_togo, var.col = addresses_with_ruca_codes$Suggested_code_rest, title = "Site practice of Togo sales by Suggested Code Rest", show.row.prc = TRUE)
# NABCA POLICIES CROSSTABS with Site-specific togo policy
sjPlot::tab_xtab(var.row = addresses_with_ruca_codes$alcohol_togo, var.col = addresses_with_ruca_codes$NABCA_spirits_bars, title = "Site practice of Togo sales by NABCA Spirits bars policy", show.row.prc = TRUE)

sjPlot::tab_xtab(var.row = addresses_with_ruca_codes$alcohol_togo, var.col = addresses_with_ruca_codes$NABCA_spirits_rest, title = "Site practice of Togo sales by NABCA spirits rest policy", show.row.prc = TRUE)
```

• I'm not surprised that there's more clustering on spirits. What is the number/percent of bars/restaurants that did not change their spirits prices?

```{r no-changers}
# Beer change
hist(addresses_with_ruca_codes$beer_price_change)

#Vodka change
hist(addresses_with_ruca_codes$vodka_price_change)

addresses_with_ruca_codes_no_change <- addresses_with_ruca_codes %>% 
  mutate(no_change_vodka = case_when(vodka_price_change == 0 ~ "1",
                                     TRUE ~ "0"),
         no_change_beer = case_when(beer_price_change == 0 ~ "1",
                                    TRUE ~ "0"))
#Differ by outlet characterisitcs, Census tract characteristics, or school characteristics 

```

I'm wondering if we can run some descriptive statistics to see if these outlets differ from those that changed their prices either in terms of outlet characteristics, school characteristics (e.g., public vs. private), or area-level characteristics.

```{r descriptive-no-changers}

```

• I am somewhat surprised to see that to-go sales were associated with higher beer prices and want to think about why that might be. Sure, beer is usually the cheapest, so there's more margin to raise prices. However, most of the changes in consumption during COVID were wine and spirits while beer consumption/sales fell, so this isn't going to be a huge revenue generator for outlets. A third possibility is that both the beer price changes and the to-go practices were concentrated in specific kinds of areas/outlets. From your exploratory analyses, did anything stand out?

• I have variables for the percent of days in 2020 that bars and restaurants were allowed to open. I'm wondering if that would explain both the to-go sales and the changes in beer price? Based on how easily I can find old records, it's possible to recode this for Feb-August or September.

Here's stuff on the covariates I was supposed to pull I'm not sure if these analyses adjust for covariates. It seems like the do from file names, but maybe the covariates were omitted from the results tables? Either way, here's the census tract covariates we discussed. Apologies for getting repeatedly sidetracked from this.

Here's the variables that I calculated: • Diversity: This is Simpson's Diversity Index. It measures the probability that two people chosen at random will be from different racial and ethnic groups. It is bounded between 0 and 1. • Disadv: This is the "disadvantage" index that Deb Furr-Holden always uses. It's calculated as ((%college-degree-or-more + %owner-occupied) -- (%poverty + %female-headed-HHs))/4. It ranges from -1 to 1; the negative values are "disadvantaged" and the positive values are advantaged. I hate the term disadvantage when ascribed to groups of people or places, which is why I keep putting it in quotes BTW.

The median HH income variable seems like the units have to be \$10,000. That would make the CT data about match what I'm assuming are county-level data in their summary tables. Here's the ACS data dictionary from 2011; I couldn't find a more recent one in a quick search. I changed variable names to be easier to understand; LMK if any aren't intuitive.
